---
header-includes: |
  \usepackage{fancyhdr}
  \pagestyle{fancy}
  \fancyhead[LO,L]{\textbf{MTH-341 A} | Spring 2025}
  \fancyhead[RO,R]{\textbf{Worksheet 9} | Pg. \thepage}
  \fancyfoot[CO,C]{}
  \usepackage{float}
  \usepackage{multirow}
output:
  bookdown::pdf_document2:
    toc: no
    number_section: no
urlcolor: blue
link-citations: yes
bibliography: ../../../references.bib
csl: ../../../mbe.csl
geometry: "left=0.50in,right=0.50in,top=0.75in,bottom=0.75in"
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

**Name:**

**Collaborators:**

\noindent\rule{\textwidth}{1pt}
**Instructions:** Worksheets are graded mostly on completion, and partially on correctness. Please write complete solutions showing explanations and key steps to the following problems, unless it says otherwise.
\noindent\rule{\textwidth}{1pt}

## Changing the Basis

### 1. Coordinate Transformation

*Changing basis* means a transformation of a vector's coordinates relative to a different basis of the same vector space. If $\displaystyle B = \{\vec{b}_1, \vec{b}_2, \cdots, \vec{b}_k\}$ is a basis and $\vec{v}$ is in standard form in $\mathbb{R}^2$, its coordinates relative to $B$ are found by solving $\displaystyle \vec{v} = c_1 \vec{b}_1 + c_2 \vec{b}_2 + \cdots + c_k \vec{b}_k.$

These coefficients form $\displaystyle \text{Rep}_{B}{\left(\vec{v}\right)} = \begin{bmatrix} c_1 \\ c_2 \\ \vdots \\ c_k \end{bmatrix}_{B}$, the representation of $\vec{v}$ with respect to $B$.

\begin{minipage}[t]{0.49\textwidth}
  \begin{itemize}
    \item[a.] Draw the standard basis basis vectors $\vec{e}_1$ and $\vec{e}_2$ in $\mathbb{R}^2$ using the blank axis below. Then, draw the vector $\displaystyle \vec{v} = \begin{bmatrix} -1 \\ -1 \end{bmatrix}$. \\
```{r fig.width=3,fig.height=3,fig.align='left'}
ggplot() +
  xlim(-2, 2) +
  ylim(-2, 2) +
  geom_hline(yintercept=0,color='black') +
  geom_vline(xintercept=0,color='black') +
  labs(x='x',y='y')+
  theme_minimal()
```
  \end{itemize}
\end{minipage}
\begin{minipage}[t]{0.49\textwidth}
  \begin{itemize}
    \item[b.] Let $\displaystyle B = \left\{\begin{bmatrix}1 \\ 1\end{bmatrix}, \begin{bmatrix}1 \\ -1\end{bmatrix}\right\}$ be a basis in $\mathbb{R}^2$. Determine $\displaystyle \text{Rep}_{B}{\left(\vec{v}\right)}$. Then, draw the vectors in $B$ and $\displaystyle \text{Rep}_{B}{\left(\vec{v}\right)}$ on the blank axis below. \\
```{r fig.width=3,fig.height=3,fig.align='right'}
ggplot() +
  xlim(-2, 2) +
  ylim(-2, 2) +
  geom_hline(yintercept=0,color='black') +
  geom_vline(xintercept=0,color='black') +
  labs(x='x',y='y')+
  theme_minimal()
```
  \end{itemize}
\end{minipage}

c. Describe your observations between the two basis vectors and the vectors $\vec{v}$ and $\displaystyle \text{Rep}_{B}{\left(\vec{v}\right)}$. Did they stretch, rotate, or both?
\vfill

\newpage

### 2. Create Your Own Basis

Recall that a set of vectors is a basis when they are both linearly independent and spans the vector space.

Let $B = \left\{\begin{bmatrix}a \\ 1\end{bmatrix}, \begin{bmatrix}1 \\ b\end{bmatrix}\right\}$ be a set of basis vectors in $\mathbb{R}^2$, where $a$ and $b$ are some real numbers.

a. Determine values $a$ and $b$ so that given the vector $\vec{v} = \begin{bmatrix}1 \\ -1\end{bmatrix}$ transforms into $\text{Rep}_{B}{\left(\vec{v}\right)}=\begin{bmatrix}-2 \\ 2\end{bmatrix}$.
\vfill

b. Confirm that the values $a$ and $b$ determined in Part (a) forms a basis $B$.
\vfill

d. Separately draw the vectors $\vec{v}$ and $\text{Rep}_{B}{\left(\vec{v}\right)}$ and their corresponding basis on the blank axis below.

\begin{minipage}[t]{0.49\textwidth}
```{r fig.width=3,fig.height=3,fig.align='center'}
ggplot() +
  xlim(-2, 2) +
  ylim(-2, 2) +
  geom_hline(yintercept=0,color='black') +
  geom_vline(xintercept=0,color='black') +
  labs(x='x',y='y')+
  theme_minimal()
```
\end{minipage}
\begin{minipage}[t]{0.49\textwidth}
```{r fig.width=3,fig.height=3,fig.align='center'}
ggplot() +
  xlim(-2, 2) +
  ylim(-2, 2) +
  geom_hline(yintercept=0,color='black') +
  geom_vline(xintercept=0,color='black') +
  labs(x='x',y='y')+
  theme_minimal()
```
\end{minipage}

e. Describe your observations between the two basis vectors and the vectors $\vec{v}$ and $\displaystyle \text{Rep}_{B}{\left(\vec{v}\right)}$. Did they stretch, rotate, or both?
\vfill

<!-- \newpage

## **References**

::: {#refs}
:::

<br> -->
